---
title: "workshop9-collab"
format: html
editor: visual
---

load libraries

```{r}

library(tidyverse) # for wrangling and plotting
library(lterdatasampler) # for data access
library(ggeffects) # for models
library(here) # for easy file pathing





```

```{r}
data(pie_crab) #reading in plum island ecosystem crab data set from lterdatasampler package
```

exploratory vis

```{r}
ggplot(aes(x = latitude,
           y = size),
       data = pie_crab) + #set up plot
  geom_point(color = "orange") +
  labs(x = "Latitude",
       y = "Size (mm)") + #update axes labels
  theme_minimal() #remove gridlines by updating theme

```


writing a model
```{r}
crab_model <- lm(size ~ latitude, #formula: response ~ predictor
                 data = pie_crab) # data form lter data sampler

# look at diagnostics ehre
par(mfrow = c(2,2)) # plot diagnostics in a 2x2 frid
plot(crab_model) #plotting the diagnostics
```

**interpret model**

```{r}
summary(crab_model)
```
slope: 0.49 +/- 0.034
y-int: -3.6+/- 1/27

For every 1 unit increase in latitude, we expect to see an increase in crab size of 0.49 +/- 0.034. 
model statistics: F(1, 390) = 208.5, p<0.0001, alpha = 0.05, R^2^ = 0.35

1. always save commit and push !!! do this often
2. TALK TO YOUR COLLABORATORS
3. Pull changes otherwise you will get a "merge conflict"

order of operations:
pull, work, save, stage, commit, close, push 

